---
title: "mesh"
author: "Peter Hahn"
date: "30 11 2018"
output: html_document
---

## moved from main script 

## prepare mesh
Extract mesh from pubmed data. Bind with article ID. Filter for variable Heading. 
## bind tf_idf
Because Mesh is hierarchical structured most documents have common terms as Dupyutren Contracture, Human, surgery, male, pathology etc. Therefore use of term frequency is not useful. Better is TF-IDF 

```{r}
        MedList = mapply(cbind, "ID"=ArticleId(fetch),Mesh(fetch),SIMPLIFY = FALSE)
        mesh_melt <-melt(MedList,id="ID")
        mesh_melt <- mesh_melt %>% filter(variable=="Heading")
        mesh_melt <- mesh_melt[,c(1,3)]
        colnames(mesh_melt) <- c("document","mesh")
        mesh_melt <- mesh_melt %>% mutate(n=1)       
        mesh_tfidf <- mesh_melt %>% bind_tf_idf(mesh,document,n)        
        
```
## Take a look at mesh
```{r}
        mesh_melt %>% group_by(mesh) %>% count() %>% arrange(-n)
```
### Take a look at mesh_tfidif
```{r}
        top_tfidf <- mesh_tfidf %>% arrange(document,desc(tf_idf)) %>% group_by(document) %>% top_n(5,tf_idf)
        #tidy(abstr_lda,matrix="gamma") %>% arrange(document,desc(gamma)) %>% group_by(document) %>% top_n(5,gamma)
```



### join lda_gamma and mesh_melt

```{r}
gamma_mesh <- inner_join(lda_gamma,top_tfidf) ##%>% filter(mesh!="Dupuytren Contracture")

        gamma_mesh %>% group_by(mesh,topic) %>% count() %>% arrange(desc(topic),desc(nn)) %>% 
                group_by(topic) %>%  top_n(10,nn) %>% ungroup() %>% filter(topic<51) %>% 
                mutate(mesh=factor(mesh, levels=rev(unique(mesh)))) %>% 
                ggplot(aes(mesh,nn))+ geom_col(show.legend = FALSE)+
                facet_wrap(~topic, ncol=4,scales="free")+ 
                coord_flip()
        

```
The top mesh terms are not specific. How many mesh terms are in each document ?

```{r}
        numb <- mesh_melt %>% group_by(document) %>% count()
        hist(numb$n)
```

### which mesh-terms are specific for each document as there are much unspecific terms
calculate TF-IDF for melt_mesh
```{r}
        
```

lda_gamma %>% group_by(document) %>% count(

)